\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xspace}
\usepackage{amsfonts}
\usepackage{algorithm2e}
\usepackage{url}
\usepackage[table,xcdraw]{xcolor}
%\usepackage{soul}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
    
\def\system{\texttt{2PLA}\xspace}
\def\dataset{\texttt{ULCER\_SET}\xspace}

\begin{document}

\title{A Two-Phase Learning Approach for the Segmentation of Dermatological Wounds*\\
\thanks{*This work was supported by FAPERJ (G. E-26/203.215/2016 and G. E-11/2018 Sediadas) and FAPESP (G. 2016/17078-0).
Corresponding authors: \textit{\{wellingtons, marcosbedo\}@id.uff.br}.}
}

%\author{\\% <-this % stops a space
%\address{}
%\thanks{$^{2}$P. M. Azevedo-Marques and A. Traina are with University of S\~{a}o Paulo.}
%\thanks{$^{3}$A. Jorge is with S\~{a}o Carlos Federal University.}
%\thanks{$^{4}$L. Santos is with Federal Institute of North of Minas Gerais.}
%\thanks{$^{5}$D. Oliveira is with IC/Fluminense Federal University - Brazil.}
%}

\author{\IEEEauthorblockN{Wellington S. Silva$^{\star}$, Daniel L. Jasbick$^{\star,\dagger}$, Rodrigo E. Wilson$^{\star}$, Paulo M. Azevedo-Marques$^{\bullet}$,\\ Agma J. M. Traina$^{\mp}$, Lucio F. D. Santos$^{\diamond}$, Ana E. S. Jorge$^{\ddagger}$, Daniel de Oliveira$^{\dagger}$, and Marcos V. N. Bedo$^{\star}$}
\IEEEauthorblockA{
$^{\star}$Fluminense Northwest Institute, Fluminense Federal University, Brazil\\
$^{\dagger}$Institute of Computing, Fluminense Federal University, Brazil\\
$^{\bullet}$Ribeir\~{a}o Preto Medical School, University of S\~{a}o Paulo, Brazil\\
$^{\mp}$Department of Computer Science, University of S\~{a}o Paulo, Brazil\\
$^{\diamond}$Federal Institute of North of Minas Gerais -- Montes Claros, Brazil\\
$^{\ddagger}$Department of Physical Therapy, Federal University of S\~{a}o Carlos, Brazil 
}}


\maketitle

\begin{abstract}
Tissue segmentation in photographs of lower limb chronic ulcers is a non-intrusive approach that supports dermatological analyses.
This paper presents \system, a method that combines supervised and unsupervised learning strategies for enhancing the segmentation of dermatological wounds.
Given an ulcer photo captured according to a fixed protocol, \system first phase performs a pixelwise classification of points of interest, whereas pre-processing filters are employed for the smoothing of image noise.
The cleaned image is further sent to the \system divide-and-conquer second phase.
It builds upon SLIC superpixel construction algorithm for dividing the lower limb into regions of interest with well-defined borders, and clusters the superpixels by taking advantage of the similarity-based DBSCAN algorithm.
We set up the phases of our method by using a real annotated set of dermatological wounds, and empirical evaluations on representative samples up to $100{,}000$ points showed a compact Multi-Layer Perceptron with Levenberg--Marquardt training algorithm (Cohen-Kappa = $.971$, Sensitivity = $.98$, and Specificity = $.98$) outperformed other classifiers as \system first phase.
Additionally, experimental trials on DBSCAN with five distance functions ($L_1$, $L_2$, $L_\infty$, Canberra, and BrayCurtis) indicated $L_1$ function provided fewer groups in comparison to the competitors, and the number of clusters was an exponential decay to the similarity ratio.
Accordingly, we used the elbow criterion for finding the $L_1$-based DBSCAN threshold as \system second phase parameterization.
We evaluated the fine-tuned setting of our method over a labeled set of ulcer images, and wounded tissues were segmented within a $.05$ Mean Absolute Error ratio.
These results illustrate the impact of learning parameters on \system as well as the method efficacy for wound segmentation.

\end{abstract}

\begin{IEEEkeywords}
Wound images, Superpixels, MLP, DBSCAN
\end{IEEEkeywords}

\input{introduction}
\input{background}
\input{method}
\input{experiments}
\input{conclusions}

\bibliographystyle{plain}
\bibliography{biblio}


%section*{References}

%Please number citations consecutively within brackets \cite{b1}. The 
%sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference 
%number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at 
%the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

%%Number footnotes separately in superscripts. Place the actual footnote at 
%the bottom of the column in which it was cited. Do not put footnotes in the 
%abstract or reference list. Use letters for table footnotes.

%Unless there are six authors or more give all authors' names; do not use 
%``et al.''. Papers that have not been published, even if they have been 
%submitted for publication, should be cited as ``unpublished'' \cite{b4}. %Papers 
%that have been accepted for publication should be cited as ``in press'' \cite{b5}. 
%Capitalize only the first word in a paper title, except for proper nouns and 
%element symbols.

%For papers published in translation journals, please give the English 
%citation first, followed by the original foreign-language citation \cite{b6}.

%\begin{thebibliography}{00}
%\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
%\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
%\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
%\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
%\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
%\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
%\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
%\end{thebibliography}

\end{document}


